### 2014-12-01 writing/F'14 paper/intro.tex
Provisional AdaDelta support; cleanup.
Big paper push; SICK data updates; hacks

% TODO: Reskim Gref paper.

@@

Since no existing model architecture can be adopted without modification
to handle entailment classification, we describe a novel method for this task
in the spirit of prior work on recursive neural networks (\ref{methods}).
In three experiments, we use this pre-specified logical grammar to
generate controlled data sets encoding the semantic relationships
between pairs of expressions and evaluate whether this model can 
learn to correctly identify the relationships between novel pairs.
In our first experiment (Section~\ref{sec:join}), we
evaluate the ability of these models to learn the core relational
algebra of natural logic from data. Our second experiment
(Section~\ref{sec:recursion}) extends this evaluation to relations
between complex recursive structures like $(a \plor b)$ and
$\plneg (\plneg a \pland \plneg b)$, and our third experiment
(Section~\ref{sec:quantifiers}) involves relations between quantified
statements like \ii{every reptile walks} and \ii{no turtle moves}.
We find that the plain RNN achieves only mixed results in all three
experiments, whereas the stronger RNTN model generalizes well in every
case, suggesting that it has in fact learned, or at least learned to
simulate, our target logical concepts. These experiments provide the 
most convincing demonstration to date of the ability of neural networks to model
semantic inferences in complex natural language sentences.

@@

% The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes. It should be noted that this is not a balanced classification problem. If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them. % TODO: How much should we say about this?

>>>>

@@

Since no existing model architecture can be adopted without modification
to handle entailment classification, we describe a novel method for this task
in the spirit of prior work on recursive neural networks (\ref{methods}).
In three experiments, we use this pre-specified logical grammar to
generate controlled data sets encoding the semantic relationships
between pairs of expressions and evaluate whether models with this architecture can 
learn to correctly identify the relationships between novel pairs.
In our first experiment (Section~\ref{sec:join}), we
evaluate the ability of these models to learn the core relational
algebra of natural logic from data. Our second experiment
(Section~\ref{sec:recursion}) extends this evaluation to relations
between complex recursive structures like $(a \plor b)$ and
$\plneg (\plneg a \pland \plneg b)$, and our third experiment
(Section~\ref{sec:quantifiers}) involves relations between quantified
statements like \ii{every reptile walks} and \ii{no turtle moves}.
We find that the plain recursive NN achieves only mixed results in all three
experiments, whereas the stronger RNTN model generalizes well in every
case, suggesting that it has in fact learned, or at least learned to
simulate, our target logical concepts. These experiments provide the 
most convincing demonstration to date of the ability of neural networks to model
semantic inferences in complex natural language sentences.

@@

% The goal of each of the three experiments that we propose is to learn classifiers that are able to classify pairs of statements from some highly constrained language into these seven classes. It should be noted that this is not a balanced classification problem. If arbitrary pairs of statements are chosen for comparison in almost any domain of natural language, the minimally informative \# relation will apply between them. %  How much should we say about this?

### 2014-12-01 writing/F'14 paper/join.tex
Big paper push; SICK data updates; hacks

% We test the model's ability to learn this behavior by creating
% artificial data sets of terms which represent sets of numbers. Since
% MacCartney and Manning's set of relations hold between sets as well as
% between sentences, we can use the underlying set structure to generate
% the all of the relations that hold between any pair of these terms, as
% in Figure \ref{lattice-figure}. We train the model defined above on a
% subset of these relations, but rather then presenting the model with a
% pair of tree-structured sentences as inputs, simply present it with
% two single terms, each of which corresponds to a single vector in the
% (randomly initialized) vocabulary matrix $V$, ensuring that the model
% has no information about the terms being compared except the relations
% between them.

@@

\paragraph{Results} 
TODO: Explain Table \ref{joinresultstable}.

\begin{table}[tp]
  \centering
  \setlength{\tabcolsep}{10pt}
  \begin{tabular}{ r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    \multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
    \midrule
    35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}\\
    \bottomrule
  \end{tabular}
  \caption{Performance on the TODO experiments. Results are reported as accuracy scores followed by macroaveraged F1 scores in parentheses.}
  \label{joinresultstable}
\end{table} % TODO: Is this worth the fair amount of space it takes up?

@@

% RNTN log: tue-j-11-6x80-hip.txt
% Train PER: 0.0021875

% RNN log: tue-j-11-6x80-r-75.txt
% Train PER: 0.047188

% and create a dataset consisting of the relations between every pair of
% sets, yielding 6400 pairs. 3200 of these pairs are then chosen as a
% test dataset, and that test dataset was further split into the 2960
% examples that can be provably derived from the test data using
% MacCartney and Manning's join table (or by the symmetry of the
% relations in about half of the cases) and the 240 that
% cannot.

% We tested a version of both the RNN model and the RNTN model on these
% data. In both cases, the models were implemented exactly as described
% in Section~\ref{methods}, but since the items being compared are
% single terms rather than full tree structures, the composition layer
% was not used, and the two models differed only in which layer function
% was used for the comparison layer. We found that the RNTN model worked
% best with 11 dimensional vector representations for the 80 sets and a
% 90 dimensional feature vector for the classifier. This model was able
% to correctly label 99.3\% of the derivable test examples, and 99.1\%
% of the remaining examples. The simpler RNN model worked best with 11
% and 75 dimensions, respectively, but was able to achieve accuracies of
% only 90.0\% and 87.\%, respectively.

% TODO: Write out results.
% TODO: Do we still care about the unprovable test examples?

These results are fairly straightforward to interpret. The RNTN model
was able to accurately encode the relations between the terms in the
geometric relations between their vectors, and was able to then use
that information to recover relations that were not overtly included
in the training data. In contrast, the RNN model was able to achieve
this behavior only incompletely. It is possible but not likely that it
could be made to find a good solution with further optimization on
different learning algorithms, or that it would do better on a larger
universe of sets, for which there would be a larger set of training
data to learn from, but the RNTN is readily able to achieve these
effects in the setting discussed here.

>>>>

@@

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{??d RNN}  & \multicolumn{2}{c}{??d RNTN}\\
    \midrule
    Provable test &52.0 &(10.3) &	-&(-)& \textbf{-} & \textbf{(-)}\\
    All test &53.8 &(10.5) &	-&(-)& \textbf{-} & \textbf{(-)}\\
    Train &53.8 &(10.5)    & -&(-)& \textbf{-} & \textbf{(-)}\\

    \bottomrule
  \end{tabular}
  \caption{Performance on the semantic relation experiments. These results and all other results on artificial data are reported as mean accuracy scores over five runs followed by mean macroaveraged F1 scores in parentheses.}
  \label{joinresultstable}
\end{table}

@@

% TODO: Do we still care about the unprovable test examples?

\paragraph{Results} 
Table \ref{joinresultstable} shows the results, which are fairly straightforward 
to interpret. The RNTN model
was able to accurately encode the relations between the terms in the
geometric relations between their vectors, and was able to then use
that information to recover relations that were not overtly included
in the training data. In contrast, the RNN model was able to achieve
this behavior only incompletely. It is possible but not likely that it
could be made to find a good solution with further optimization on
different learning algorithms, or that it would do better on a larger
universe of sets, for which there would be a larger set of training
data to learn from, but the RNTN is readily able to achieve these
effects in the setting discussed here.

### 2014-12-01 writing/F'14 paper/methods.tex
Big paper push; SICK data updates; hacks

We use the model architecture depicted in Figure~\ref{sample-figure}. 
The two phrases being compared are processed separately using a pair 
of recursive networks that share a single set of parameters. 
The resulting vectors are fed into a separate comparison
layer that is meant to generate a feature vector capturing the
relation between the two phrases. The output of this layer is then
given to a softmax classifier, which in turn produces a hypothesized
distribution over the seven relations represented in Table~\ref{b-table}.

For a composition layer, we evaluate models with both the plain neural
network layer function \eqref{rnn} and a more powerful extension, the RNTN layer function
\eqref{rntn} proposed in \newcite{chen2013learning}. A $\tanh$ nonlinearity is applied to the output of the layer function.
%
\begin{gather} \label{rnn}
\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\ % TODO: Add column vectors?
\label{rntn}
\vec{y}_{\textit{RNTN}} = \vec{y}_{\textit{RNN}} + f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather} 
%
Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
representations for the left and right children of the node, and
$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
them by an $n \times 2n$ matrix of learned weights, adds a bias $\vec{b}$, and applies the
element-wise non-linearity to the resulting vector. The RNTN adds a learned third-order tensor $\mathbf{T}$, dimension $n \times n \times n$, modeling
multiplicative interactions between the child vectors.

% TODO: Kill ReLU?

@@

% Removed 'minibatch' -> For two of the experiments, we use minibatches
% of size 1, which doesn't really count.
We train the model using stochastic gradient descent (SGD)
with learning rates computed using AdaGrad \cite{duchi2011adaptive}.
The parameters (including the vocabulary) are initialized randomly.
Because the classes are not balanced in general, we report performance
using both accuracy and macroaveraged F1, since the latter emphasizes
 performance on infrequent classes. We compute macroaveraged F1 
as the harmonic mean of mean precision and mean recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined. Source code and generated data will be released
after the review period.

% TODO: Describe mean-of-five scheme for results reporting.

>>>>

We use the model architecture depicted in Figure~\ref{sample-figure}. 
The two phrases being compared are processed separately using a pair 
of recursive networks that share a single set of parameters. 
The resulting vectors are fed into a separate comparison
layer (fixed at 75 dimensions) that is meant to generate a feature vector capturing the
relation between the two phrases. The output of this layer is then
given to a softmax classifier, which in turn produces a hypothesized
distribution over the seven relations represented in Table~\ref{b-table}.

For a composition layer, we evaluate models with both the plain neural
network layer function \eqref{rnn} and a more powerful extension, the RNTN layer function
\eqref{rntn} proposed in \newcite{chen2013learning}. A $\tanh$ nonlinearity is applied to the output of the layer function.
%
\begin{gather} \label{rnn}
\vec{y}_{\textit{RNN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) \\
\label{rntn}
\vec{y}_{\textit{RNTN}} = \vec{y}_{\textit{RNN}} + f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather} 
%
Here, $\vec{x}^{(l)}$ and $\vec{x}^{(r)}$ are the column vector
representations for the left and right children of the node, and
$\vec{y}$ is the node's output.  The RNN concatenates them, multiplies
them by an $n \times 2n$ matrix of learned weights, adds a bias $\vec{b}$, and applies the
element-wise non-linearity to the resulting vector. The RNTN adds a learned third-order tensor 
$\mathbf{T}$, dimension $n \times n \times n$, modeling
multiplicative interactions between the child vectors.

@@

% Removed 'minibatch' -> For two of the experiments, we use minibatches
% of size 1, which doesn't really count.
We train the model using stochastic gradient descent (SGD)
with learning rates computed using ADADELTA \cite{zeiler2012adadelta}.
The parameters (including the vocabulary) are initialized randomly.
Because the classes are not balanced in general, we report performance
using both accuracy and macroaveraged F1, since the latter emphasizes
 performance on infrequent classes. We compute macroaveraged F1 
as the harmonic mean of mean precision and mean recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined. Source code and generated data will be released
after the review period.

### 2014-12-01 writing/F'14 paper/quantifiers.tex
Provisional AdaDelta support; cleanup.
Big paper push; SICK data updates; hacks

% TODO: We generate clean data from a grammar.

@@

% TODO: Rewrite experimental setup.
We evaluate the model using two experimental settings. In the simpler
setting, \textsc{all split}, we randomly sample 85\% of the data and evaluate on the
remaining 15\%. In this setting, the model is being asked to learn a
complete reasoning system for the limited language and logic presented
in the training data, but it is not being asked to generalize to test
examples that are substantially different from those it was trained
on. Crucially, though, to succeed on this task, the model must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
\eqref{p1} and \eqref{p2} in training and, from that information,
determine \eqref{p3}.
%
% do not allow a blank line --- adds too much space
%
\begin{gather}
  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
  \\
  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
  \\
  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
\end{gather}
%
% do not allow a blank line --- adds too much space
%
In this configuration the model ...

\begin{table}[tp]
  \centering
  \setlength{\tabcolsep}{10pt}
  \begin{tabular}{ r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    \multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
    \midrule
    35.4 &(7.5) &	67.4&(56.5)& \textbf{100.0} & \textbf{(100.0)}\\
    \bottomrule
  \end{tabular}
  \caption{Performance on the quantifier experiments.}
  \label{qresultstable}
\end{table}

\paragraph{Results} % TODO: Rewrite!
The results for these experiments are shown in
Table~\ref{qresultstable}. We compare the results to a most frequent
class baseline, which reflects the frequency in the test data of the
most frequent class in the training data, $\natind$.  After some
cross-validation, we chose 16- and 20-dimensional representations for
the RNN and RNTN, respectively, and 75-dimensional feature vectors for
the classifier. Training
accuracy was near 80\% for the each of the five RNN experiments, and 
near 98\% for each of the five RNTN experiments.

Discussion discussion. 
Discussion discussion. 
Discussion discussion. 
Discussion discussion. 
Discussion discussion. 
Discussion discussion. 
Discussion discussion. 
Discussion discussion. 
Discussion discussion.

>>>>

@@

% TODO: Rewrite experimental setup.
We evaluate the model by randomly sampling 20\% of the data to
test on and training on the remainder. In this setting, the model is being asked to learn a
complete reasoning system for the limited language and logic presented
in the training data, but it is not being asked to generalize to test
examples that are substantially different from those it was trained
on. Crucially, though, to succeed on this task, the model must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
\eqref{p1} and \eqref{p2} in training and, from that information,
determine \eqref{p3}.
%
% do not allow a blank line --- adds too much space
%
\begin{gather}
  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
  \\
  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
  \\
  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
\end{gather}
%
% do not allow a blank line --- adds too much space
%
\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~&\multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{??d RNN}  & \multicolumn{2}{c}{??d RNTN}\\
    \midrule
    Test & 35.4 &(7.5) &	99.1&(99.2)& \textbf{-} & \textbf{(-)}\\
    Train & 35.4 &(7.5) &	-&	-&	-&	-\\	
    \bottomrule
  \end{tabular}
  \caption{Performance on the quantifier experiments.}
  \label{qresultstable}
\end{table} 
%
% do not allow a blank line --- adds too much space
%
\paragraph{Results} The results, reported in Figure \ref{qresultstable}, show that both models are able to learn to generalize the underlying logic well.

% TODO: More to say?

### 2014-12-01 writing/F'14 paper/sick.tex
Big paper push; SICK data updates; hacks

What SICK is. \cite{marelli2014sick}
\\ \\ \\ \\

\begin{table*}[htp]
  \centering
  \begin{tabular}{lcl}
    \toprule
    (all turtles) walk      &    $\natind$  & (less-than-three turtles) walk\\
    (three warthogs) move  &        $\natneg$ & (less-than-three warthogs) move\\
    (all mammals) (not move) &         $\natalt$ & (some warthogs) swim\\
    (some (not pets)) swim     &     $\natcov$  & (all (not pets)) (not move)\\
    \bottomrule
  \end{tabular}
  \caption{\label{examplesofsickdata}Examples of successful classifications on SICK.}
\end{table*}

We initialized our word embeddings using precomputed TODO dimensional vectors from 
(TODO: Where?). Before any embedding is used as an input to a recursive layer, 
it is passed through an additional $\tanh$ neural network layer that transforms
the embedding from the embedding
dimension to the dimension of the recursive layer.

We supplemented the roughly 5000 SICK training examples\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with training data from (CITE CITE CITE CITE CITE CITE CITE CITE CITE CITE CITE). We trained a single model on data from all sources, but used a separate set of softmax parameters for classifying into the labels from each source.

\begin{table}[tp]
  \centering
  \setlength{\tabcolsep}{10pt}
  \begin{tabular}{ r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    \multicolumn{2}{c}{$\natind$ only} & \multicolumn{2}{c}{16 dim RNN}  & \multicolumn{2}{c}{20 dim RNTN}\\
    \midrule
    56.7 &(...) &	...&(...)& \textbf{...} & \textbf{(...)}\\
    \bottomrule
  \end{tabular}
  \caption{Performance on the SICK.}
  \label{sresultstable}
\end{table}

Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}. 
Explain Figure \ref{examplesofsickdata}.

% TODO: Last runs, train on dev.

>>>>

The tree-pair model architecture that we use is novel, and though the underlying recursive approach has been validated elsewhere, there is no guarantee that this architecture is suitable for handling inference over real natural language. To ensure that this is the case, we also train a version of our model on an existing competition dataset, the ICK textual entailment challenge \cite{marelli2014sick}. The corpus consists of about 10k natural language sentence pairs (generated using data expansion from a smaller set), labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}. At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest hand labeled entailment corpus that we are aware of.

\begin{table*}[htp]
  \centering
  \begin{tabular}{l}
    \toprule
example example example example example example example example example example example \\
example example example example example example example example example example example \\
ENTAILMENT A skilled person is riding a bicycle on one wheel	A person is riding the bicycle on one wheel\\
CONTRADICTION The player is missing the basket and a crowd is in background	The player is dunking the basketball into the net and a crowd is in background\\
    \bottomrule
  \end{tabular}
  \caption{\label{examplesofsickdata}Examples of successful classifications on SICK.}
\end{table*}

We initialized our word embeddings using the TODO dimensional vectors provided with 
GloVe \cite{pennington2014glove}. Before any embedding is used as an input to a recursive layer, 
it is passed through an additional $\tanh$ neural network layer that transforms
the embedding from the embedding dimension to the dimension of the recursive layer.

We supplemented the roughly 5000 SICK training examples\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set.} with TODO examples of approximate entailment data from the Denotation Graph project \cite{hodoshimage}. We parsed all data with the Stanford PCFG Parser v. 3.3.1 \cite{klein2003accurate}. To accelerate convergence, we collapsed subtrees that were identical across both sentences in a pair down to a single head word. In order to improve regularization on the noisier data, we used dropout \cite{hinton2012improving} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%). We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source.

\begin{table}[tp]
  \centering \small
  \begin{tabular}{ l r@{ \ } r@{ \ } r@{ \ } }
    \toprule
    ~&\multicolumn{1}{c}{$\natind$ only} & \multicolumn{1}{c}{??d RNN}  & \multicolumn{1}{c}{??d RNTN}\\
    \midrule
    Test & 56.7 &	...& \textbf{...} \\
    Train &- &- &-  \\
    \bottomrule
  \end{tabular}
  \caption{Performance on SICK.}
  \label{sresultstable}
\end{table}

Despite the small amount of high quality training data available, it is possible to train our model to perform competitively on textual entailment. Our performance did not reach the state of the art (84.6\%), but exceeded the median submission to the competition (77.06\%) in a field including many models which used sophisticated hand-engineered features specific to the specific version of the entailment task at hand.

Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}. 
Explain Table \ref{examplesofsickdata}.
